Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes

### 2021-03-23 code/trash_can.Rmd
fixed nomenclature issue

>>>>

-------------------------------------
# Old

## 3.6 Genus-level niche overlap ~ tongue.pair + elevation
### 3.6.1 Model
```{r}
niche_gam10 <- gam(mean.niche.overlap ~
                     tongue.pair +
                     s(elev.mean, by = tongue.pair, k = 8, m = 2),
                   method = "REML",
                   family = betar(eps = 0.001),
                   select = TRUE,
                   data = niche_overlap_tongues) %>% getViz()

check.gamViz(niche_gam10)
summary(niche_gam10)
print(plot(niche_gam10, allTerms = TRUE), pages = 1)
```

### 3.6.2 Visualize
```{r}
# Get predictions using tidymv
niche_gam10_pred <- get_gam_predictions(niche_gam10, elev.mean) %>%
  as_tibble() %>%
  mutate(mean.niche.overlap_trans = inv_logit_scaled(mean.niche.overlap),
         CI_upper_trans = inv_logit_scaled(CI_upper),
         CI_lower_trans = inv_logit_scaled(CI_lower))

# Plot smooths
ggplot(niche_gam10_pred, aes(elev.mean, mean.niche.overlap_trans, color = tongue.pair)) +
  geom_ribbon(aes(elev.mean, mean.niche.overlap_trans,
                  ymax = CI_upper_trans, ymin = CI_lower_trans,
                  fill = tongue.pair),
              alpha = 0.3, inherit.aes = FALSE) +
  geom_line() +
  ylab("Niche overlap") +
  xlab("Elevation") +
  theme_light()

ggsave(file = "../output/niche_gam10_plot1.pdf", width = 5, height = 4)
ggsave(file = "../output/niche_gam10_plot1.png", width = 5, height = 4)
```

### 3.5.1 Model
```{r}
brm_niche00 <- brm(niche.overlap ~ 1 +
                     level + # fixed effect of level 
                     (1 | site) + # varying intercept for site
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_00", # local
                   #file = "./brm_niche_00", # HPC
                   iter = 2000,
                   chains = 4,
                   data = niche_pool)

pp_check(brm_niche00)
bayes_R2(brm_niche00)

get_variables(brm_niche00)
conditional_effects(brm_niche00)
mcmc_plot(brm_niche00, pars = c("^r_sp.pair\\[mend*"))
mcmc_plot(brm_niche00, pars = c("^r_site*"))
```

### 3.5.2 Vizualize
```{r}
brm_niche00_plotframe1 <- brm_niche00 %>%
  spread_draws(b_Intercept, b_levelgenus, b_levelmorphotype) %>%
  mutate(b_levelgenus = b_Intercept + b_levelgenus,
         b_levelmorphotype = b_Intercept + b_levelmorphotype) %>%
  pivot_longer(-c(.chain, .iteration, .draw), names_to = "level", values_to = "value") %>%
  group_by(level) %>%
  point_interval(value, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(value_trans = inv_logit_scaled(value), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>%
  mutate(level = case_when(
    level == "b_Intercept" ~ "family",
    level == "b_levelmorphotype" ~ "morphotype",
    level == "b_levelgenus" ~ "genus"
  )) %>%
  mutate(level = factor(level, levels = c("genus", "family", "morphotype")))

ggplot(brm_niche00_plotframe1, aes(value_trans, level, 
                                 xmin = .lower_trans, xmax = .upper_trans)) +
  geom_pointinterval() +
  theme_light(12) +
  labs(x = "Niche overlap", y= "Level") +
  guides(fill = FALSE) +
  scale_x_continuous(breaks = seq(0.25, 0.5, by = 0.05))

ggsave(file = "../output/brm_niche00_plot1.pdf", width = 4.5, height = 4)
ggsave(file = "../output/brm_niche00_plot1.png", width = 4.5, height = 4)
```

## 3.6 Genus niche overlap ~ tongue
### 3.6.1 Model
```{r}
brm_niche10 <- brm(niche.overlap ~ 1 +
                     (1 | site) + # varying intercept for site
                     (1 | tongue.pair) +
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_10", # local
                   #file = "./brm_niche_10", # HPC
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_pool, level == "genus"))

pp_check(brm_niche10)
bayes_R2(brm_niche10)

get_variables(brm_niche10)
mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
mcmc_plot(brm_niche10, pars = c("^r_site*"))
```

### 3.6.2 Vizualize
```{r}
brm_niche10_plotframe1 <- brm_niche10 %>%
  spread_draws(b_Intercept, r_tongue.pair[condition, term]) %>%
  mutate(condition_mean = b_Intercept + r_tongue.pair) %>%
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>%
  mutate(condition = factor(condition, levels = c("short:short", "med:med", "long:long",
                                          "short:med", "short:long", "med:long")))

ggplot(brm_niche10_plotframe1, aes(condition_mean_trans, condition, 
                                 xmin = .lower_trans, xmax = .upper_trans)) +
  geom_pointinterval() +
  theme_light(12) +
  labs(x = "Niche overlap", y= "Tongue pair") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_niche10_plot1.pdf", width = 4.5, height = 4)
ggsave(file = "../output/brm_niche10_plot1.png", width = 4.5, height = 4)
```

## 3.7 Genus niche overlap ~ sp.pair
### 3.7.1 Model
```{r eval=FALSE, include=FALSE}
brm_niche20 <- brm(niche.overlap ~ 1 +
                     (1 | site) + # varying intercept for site
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_20", # local
                   #file = "./brm_niche_20", # HPC
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_overlap, level == "genus"))

pp_check(brm_niche20)
bayes_R2(brm_niche20)

get_variables(brm_niche20)
conditional_effects(brm_niche20)
mcmc_plot(brm_niche20, pars = c("^r_sp.pair"))
mcmc_plot(brm_niche20, pars = c("^r_site*"))
```

### 3.7.2 Visualize
```{r eval=FALSE, include=FALSE}
brm_niche20_plotframe1 <- brm_niche20 %>%
  spread_draws(b_Intercept, r_sp.pair[condition, term]) %>%
  mutate(condition_mean = b_Intercept + r_sp.pair) %>%
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>%
  mutate(condition = factor(condition)) %>%
  left_join(niche_pool, by = c("condition" = "sp.pair")) %>%
  select(condition, condition_mean_trans, .lower_trans, .upper_trans, tongue.pair) %>%
  distinct() %>%
  mutate(tongue.pair = factor(tongue.pair, levels = c("short:short", "med:med", "long:long",
                                                      "short:med", "short:long", "med:long")))

ggplot(brm_niche20_plotframe1, aes(condition_mean_trans, condition, 
                                 xmin = .lower_trans, xmax = .upper_trans)) +
  geom_pointinterval() +
  theme_light(11) +
  labs(x = "Niche overlap", y= "Tongue pair") +
  guides(fill = FALSE) +
 # facet_wrap(~tongue.pair, ncol = 1, scales = "free_y") +
  facet_wrap(~tongue.pair, scales = "free_y")

ggsave(file = "../output/brm_niche20_plot1.pdf", width = 4.5, height = 4)
ggsave(file = "../output/brm_niche20_plot1.png", width = 7, height = 10)
```

## 3.8 Niche overlap ~ elevation
### 3.8.1 Models
```{r}
ggplot(niche_overlap_sppair, aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

ggplot(filter(niche_overlap_sppair, 
              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") & 
                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

ggplot(filter(niche_overlap_sppair, 
              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf",
                         "mont", "muci", "pyre", "mend", "gers") & 
                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
                           "mont", "muci", "pyre", "mend", "gers")), 
       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

ggplot(filter(niche_overlap_sppair,
              sp1 %in% c("pasc", "prat", "soro", "telu", "wurf") | 
                sp2 %in% c("pasc", "prat", "soro", "telu", "wurf")), 
       aes(elev.mean, mean.niche.overlap, color = sp.pair)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = FALSE) +
  facet_wrap(~sp.pair) +
  ylim(c(0,1))

# Pairwise
gen_gam_niche.I <- bam(mean.niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(site, bs = "re") + # since I have measures repeated over species pairs within each elevation level
                            s(elev.mean, by = sp.pair, k = 8, m = 2),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_overlap_sppair, level == "genus" &
                                          sp1 %in% c("pasc", "prat", "soro", "telu", "wurf", # dominant species
                                                        "mont", "muci", "pyre", "mend") & # high elevation specialists
                                          sp2 %in% c("pasc", "prat", "soro", "telu", "wurf",
                                                        "mont", "muci", "pyre", "mend"))) %>% getViz()

check.gamViz(gen_gam_niche.I)
summary(gen_gam_niche.I)
plot(sm(gen_gam_niche.I, 1))
plot(sm(gen_gam_niche.I, 11))
plot(sm(gen_gam_niche.I, 28))
plot(sm(gen_gam_niche.I, 33))
plot(sm(gen_gam_niche.I, 37))

# Community mean
ggplot(niche_pool_sitemean, aes(elev.mean, mean.niche.overlap, color = level)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

### 2021-03-23 ms_2/ms_2_v2.Rmd
fixed nomenclature issue

Species coexistence is constrained in part by competition over shared food sources. Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities. We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps. Using these data, we first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation. We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes. This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths. Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning of floral taxa within morphotypes. This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species. Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient. If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations in cases where floral resources are limiting.

@@

<!-- Don't forget: I need to fix the Gentiana/Gentianella/Gentianopsis nomenclature problem and rerun all the models before I finalize the paper. This won't change anything about the interpretation of the  analysis, but it needs to be done. It might make Gentiana show up in the PCoA. -->

@@

Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy]. In this role, however, they have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy]. To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie]. In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that factors other than food availability --- such as abiotic stressors or predator/pathogen pressure --- limit populations.

@@

All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Chesson2000-aw]. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth as poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g. @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. This association between tongue length and floral niche, though, is imperfect, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g. @Ranta1980-ok; @Goulson2008-oe].

Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition. It has been well-demonstrated that niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g. @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi]. Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk]. This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition --- which cannot be inferred from niche overlap --- is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.

@@

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz. While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers, which we pooled. During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses. In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^. Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

@@

We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. Five species --- *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites. *B. hortorum*, *B. jonellus*, and *B. psithyrus* (species-group) were widespread but at lower abundance. *B. monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m. The remaining species were relatively rare in our study system. *B. mendax*, *B. mucidus*, and *B. pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci]. *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius* were sparsely distributed across mid-elevation sites. *B. humilis* was recorded only three times and only in one year, so we omitted it from all analyses.

@@

```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation. Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show the standard errors of the mean")}
knitr::include_graphics("../output/niche_gam00_plot1.png")
```

```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show standard errors of the mean")}
knitr::include_graphics("../output/niche_gam20_plot1.png")
```

@@

We thank D.J. McNeil for helpful conversations during data analysis.

>>>>

Species coexistence is constrained in part by competition over shared food sources. Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to reshuffle competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities. We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps. Using these data, we first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation. We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes. This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths. Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning within floral morphotypes. This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species. Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient. If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations in cases where floral resources are limiting.

@@

@@

Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy]. In this role, however, they have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy]. To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie]. In some cases, competition for food may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr], implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure.

@@

All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Chesson2000-aw]. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral selection in bumble bee communities [e.g. @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. This association between tongue length and floral niche, though, is imperfect, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g. @Ranta1980-ok; @Goulson2008-oe].

Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition. It has been well-demonstrated that niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g. @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi]. Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk]. This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.

@@

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz. While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers, which we pooled. In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^. Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

@@

We recorded a total of 12,918 bumble- bee-flower interactions over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. Five species --- *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites. *B. hortorum*, *B. jonellus*, and *B. psithyrus* (species-group) were widespread but at lower abundance. *B. monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m. The remaining species were relatively rare in our study system. *B. mendax*, *B. mucidus*, and *B. pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci]. *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius* were sparsely distributed across mid-elevation sites. *B. hypnorum* and *B. humilis* were recorded only three times and only in one year, so we omitted it from all analyses.

@@

```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation. Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\% confidence intervals.")}
knitr::include_graphics("../output/niche_gam_00_plot1.png")
```

```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\% confidence intervals.")}
knitr::include_graphics("../output/niche_gam_20_plot1.png")
```

@@

We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment for expediting Bayesian model fitting.
